Hybrid recurrent model for regime switching nonstationary time series

### 2021-11-24 iclr2022/hybridmodel_iclr.tex
Corrections appendix
A few updates

%During the past decades, forecasting applications have taken a totally new proportion. 
In many fields, single or few time series have become thousands of sequences with complex statistical structures. In this new context, classical time series specific statistical models show limitations when dealing with numerous heterogeneous data. Recurrent neural networks and recent sequence to sequence deep learning architectures offer very appealing numerical alternatives thanks to their capability of leveraging any kind of heterogeneous multivariate data, see for instance \citep{ hochreiter1997long,vaswani2017attention, salinas2020deepar, li2019enhancing, lim2019temporal}. %In the past decades, with the impressive results of deep neural networks (DNN) in vision or language, some time series forecasting applications have been proposed. 
Significant improvements have been achieved with Reccurent Neural Network and its Long Short-Term Memory (LSTM) or Gated Recurrent Unit (GRU)  declinations \citep{8614252}. 
A good example is the DeepAR model \citep{salinas2020deepar}. Based on a LSTM architecture, it can predict accurate probability distributions of the future of a sequence. More recently, applications using the Transformer model have been proposed  \citep{li2019enhancing}. A direct concurrent of the DeepAR model can be found with the Temporal Fusion Transformers (TFT) approach \citep{lim2019temporal}.  Unfortunately, they suffer from two main weaknesses. Firstly, many of them are "black-boxes"  as the final forecast usually does not come with a statistical guarantee  although a few recent works focused on measuring uncertainty in recurrent neural networks, see  \cite{martin2020monte}. Secondly, without a fine prepossessing and well chosen hyper-parameters, these methods could lead to poor results and are often outperformed by traditional statistical parametric models, see \cite{makridakis2018m4}.

@@

Recurrent neural network are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining DNN architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \citep{zhang2003time,jianwei2019novel,bandara2020lstm}. The approach providing the most striking results was proposed in  \citep{smyl2020hybrid} in the context of the M4 forecasting competition \citep{makridakis2020m4}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model is introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network is trained on the entire dataset to correct errors of the previous exponential smoothing models. %With this process, the hybrid approach succeeds at mixing strengths of both model families: robustness of the time series specific parametric models and the power of neural network at learn complex dynamics on large dataset. 
%However, the hybrid model has not been designed to deal with external signals even if it include a neural network.

Following this work, we present in this paper a new hyrbid recurrent model for time series forecasting with inclusion of external weak signals. %We design a more general hybrid framework including the original multiplicative model and a new additive one. 
Our approach enriches existing hybrid models by incorporating an unobserved regime switching feature to account for the need neural network based correction. In order to leverage external signals, we modify the hybrid framework to allow the regime shift and the  global error-corrector neural network to deal with external data. By adding our fashion external weak signals in our proposed framework, we show the real potential of the hybrid approach: a global neural network, able to leverage large amount of data, deal with any kind of external weak signals, learn context and finally correct weaknesses and errors of any parametric models.

@@

During the M4 competition, the hybrid model of \cite{smyl2020hybrid} is based on a multiplicative exponential smoothing model as the time-series-specific model. However, on sporadic time series, this choice could lead to poor result and instability. In our approach, we decided to go deeper and provide a general framework able to deal with any kind of per-time-series models. In Section~\ref{sec:exp}, we present two versions of our framework. The first one use an  exponential smoothing as reference  similar to \cite{smyl2020hybrid} and the second one use a TBATS model \cite{doi:10.1198/jasa.2011.tm09771} which provides better results. The main complexity is in the next stage: training a global RNN to correct the per-time-series statistical models when the statistical models provide poor performance.
% Indeed, depending on time series, some family of parametric models can achieve really heterogeneous level of accuracy. 
%For all $1\leqslant n \leqslant N$, let $\Theta^n = \{\alpha^n, \gamma^n, \lvl^n_0, \{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}\}$ be the parameters of the $n$-th sequence with $\lsmooth^n$ and $\ssmooth^n$ the positive smoothing parameters, $\lvl^n_0$ the first level term and $\{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}$ the $m$ first seasonality terms. Given the sequence $\fullts^n$ and the parameters $\Theta^n$ , the two components  $\fulllvl^n$ and $\fullseas^n$ are computed as follow :
%We initialize the first level with $\lvl^n_{0} = \sum_{t = 1}^m \ts^n_{t}/m$ \textcolor{red}{issue with the time index ?} and, for all $1\leqslant n \leqslant N$ and $1\leqslant t \leqslant T$,
%\begin{align*}
    %\lvl^n_{t} &= \lsmooth^n (\ts^n_t - \seas^n_{t-m}) + (1- \lsmooth^n)\lvl^n_{t-1}\,, \\
    %\seas^n_t &= \ssmooth^n (\ts^n_t - \lvl^n_{t-1}) + (1 - \ssmooth^n)\seas^n_{t-1}\,. 
%\end{align*}

@@

To estimate this model, we introduce our new regime swithing error-corrector framework. We train two global RNN on all sequences, a classifier ($\classifier$) and a predictor ($\predictor$). We train the classifier to estimate the law of the Bernoulli random variable $\hiddenregime^n_T$ we train the predictor to predict $\{\remainder^{n}_{T+i}\}_{1\leqslant i\leqslant \lag}$. However, these two tasks require a thorough data pre-processing and with our collection of parametric models, we propose first to normalize and deseasonalize the RNNs inputs.

@@

\begin{align}
\mathbb{Q}^n(\rnnparam|\rnnparam^{(t)}) &= \mathbb{E}\left[\nabla_\theta \log p_\theta(\ts^n_{T+1:T+h},U_T^n|\ts^n_{1:T})\middle|\ts^n_{1:T+h},\rnnparam^{(t)}\right]\ \\
& = \sum_{U_T^n} p(U_T^n|\ts^n_{1:T+h},\rnnparam^{(t)}) \times \log p_\theta(\ts^n_{T+1:T+h},U_T^n|\ts^n_{1:T}) \\
& = \sum_{U_T^n} p(U_T^n|\ts^n_{1:T+h},\rnnparam^{(t)}) \times \log[p_\theta(U_T^n|\ts^n_{1:T})p_\theta(\ts^n_{T+1:T+h}|U_T^n,\ts^n_{1:T})]
\end{align}

>>>>

%During the past decades, forecasting applications have taken a totally new proportion. 
In many fields, single or few time series have become thousands of sequences with complex statistical structures. In this new context, classical time series specific statistical models show limitations when dealing with numerous heterogeneous data. Recurrent neural networks and recent sequence to sequence deep learning architectures offer very appealing numerical alternatives thanks to their capability of leveraging any kind of heterogeneous multivariate data, see for instance \citep{ hochreiter1997long,vaswani2017attention, salinas2020deepar, li2019enhancing, lim2019temporal}. %In the past decades, with the impressive results of deep neural networks (DNN) in vision or language, some time series forecasting applications have been proposed. 
Significant improvements have been achieved with Reccurent Neural Network and its Long Short-Term Memory (LSTM) or Gated Recurrent Unit (GRU)  declinations \citep{8614252}. 
A good example is the DeepAR model \citep{salinas2020deepar}. Based on a LSTM architecture, it can predict accurate probability distributions of the future of a sequence. More recently, applications using the Transformer model have been proposed  \citep{li2019enhancing}. A direct concurrent of the DeepAR model can be found with the Temporal Fusion Transformers (TFT) approach \citep{lim2019temporal}.  Unfortunately, they suffer from two main weaknesses. Firstly, many of them are "black-boxes"  as the final forecast usually does not come with a statistical guarantee  although a few recent works focused on measuring uncertainty in recurrent neural networks, see  \cite{martin2020monte}. Secondly, without a fine prepossessing and well chosen hyper-parameters, these methods could lead to poor results and are often outperformed by traditional statistical models, see \cite{makridakis2018m4}.

@@

Recurrent neural networks are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining DNN architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \citep{zhang2003time,jianwei2019novel,bandara2020lstm}. The approach providing the most striking results was proposed in  \citep{smyl2020hybrid} in the context of the M4 forecasting competition \citep{makridakis2020m4}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model is introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network is trained on the entire dataset to correct errors of the previous exponential smoothing models. %With this process, the hybrid approach succeeds at mixing strengths of both model families: robustness of the time series specific parametric models and the power of neural network at learn complex dynamics on large dataset. 
%However, the hybrid model has not been designed to deal with external signals even if it include a neural network.

Following this work, we present in this paper a new hyrbid recurrent model for time series forecasting with inclusion of external weak signals. %We design a more general hybrid framework including the original multiplicative model and a new additive one. 
Our approach enriches existing hybrid models by incorporating an unobserved regime switching feature to account for the need of a neural network based correction. In order to leverage external signals, we modify the hybrid framework to allow the regime shift and the  global error-corrector neural network to deal with external data. By adding our fashion external weak signals in our proposed framework, we show the real potential of the hybrid approach: a global neural network, able to leverage large amount of data, deal with any kind of external weak signals, learn context and finally correct weaknesses and errors of any parametric models.

@@

During the M4 competition, the hybrid model of \cite{smyl2020hybrid} is based on a multiplicative exponential smoothing model as the time-series-specific model. However, on sporadic time series, this choice could lead to poor result and instability. In our approach, we decided to go deeper and provide a general framework able to deal with any kind of per-time-series models. In Section~\ref{sec:exp}, we present two versions of our framework. The first one use an  exponential smoothing as reference  similar to \cite{smyl2020hybrid} and the second one use a TBATS model \cite{doi:10.1198/jasa.2011.tm09771} which provides better results. The main complexity is in the next stage: training a global RNN to correct the per-time-series statistical models when the statistical models provide poor performance. \textcolor{red}{Plutot dire ici que l'on souhaite corriger des erreurs specifiques: et introduire les types de trends mal predites que l'on souhaite corriger.}
% Indeed, depending on time series, some family of parametric models can achieve really heterogeneous level of accuracy. 
%For all $1\leqslant n \leqslant N$, let $\Theta^n = \{\alpha^n, \gamma^n, \lvl^n_0, \{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}\}$ be the parameters of the $n$-th sequence with $\lsmooth^n$ and $\ssmooth^n$ the positive smoothing parameters, $\lvl^n_0$ the first level term and $\{\seas_t^n\}_{-m+1\leqslant t \leqslant 0}$ the $m$ first seasonality terms. Given the sequence $\fullts^n$ and the parameters $\Theta^n$ , the two components  $\fulllvl^n$ and $\fullseas^n$ are computed as follow :
%We initialize the first level with $\lvl^n_{0} = \sum_{t = 1}^m \ts^n_{t}/m$ \textcolor{red}{issue with the time index ?} and, for all $1\leqslant n \leqslant N$ and $1\leqslant t \leqslant T$,
%\begin{align*}
    %\lvl^n_{t} &= \lsmooth^n (\ts^n_t - \seas^n_{t-m}) + (1- \lsmooth^n)\lvl^n_{t-1}\,, \\
    %\seas^n_t &= \ssmooth^n (\ts^n_t - \lvl^n_{t-1}) + (1 - \ssmooth^n)\seas^n_{t-1}\,. 
%\end{align*}

@@

To estimate this model, we introduce our new regime swithing error-corrector framework. We train two global RNN on all sequences, a classifier ($\classifier$) and a predictor ($\predictor$). We train the classifier to estimate the law of the Bernoulli random variable $\hiddenregime^n_T$ and we train the predictor to predict $\{\remainder^{n}_{T+i}\}_{1\leqslant i\leqslant \lag}$. However, these two tasks require a thorough data pre-processing and with our collection of parametric models, we propose first to normalize and deseasonalize the RNNs inputs.

@@

\begin{align}
\mathbb{Q}^n(\rnnparam|\rnnparam^{(t)}) &= \mathbb{E}\left[\nabla_\theta \log p_\theta(\ts^n_{T+1:T+h},U_T^n|\ts^n_{1:T})\middle|\ts^n_{1:T+h},\rnnparam^{(t)}\right]\ \\
& = \sum_{j} p(U_T^n = j|\ts^n_{1:T+h},\rnnparam^{(t)}) \times \textcolor{red}{\nabla_\theta} \log p_\theta(\ts^n_{T+1:T+h},U_T^n = j|\ts^n_{1:T}) \\
& = \sum_{j} p(U_T^n = j|\ts^n_{1:T+h},\rnnparam^{(t)}) \times \textcolor{red}{\nabla_\theta}\log[p_\theta(U_T^n = j|\ts^n_{1:T})p_\theta(\ts^n_{T+1:T+h}|U_T^n = j,\ts^n_{1:T})]
\end{align}
